---
title: "FIESTA - Small Area Estimators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FIESTA - Small Area Estimators}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Small Area (SA) module overview

`FIESTA`'s Small Area (SA) module was set up as a platform to integrate with current Small Area Estimators available on CRAN including the `JoSAE` (Breidenbach 2015), `sae` (Molina and Marhuenda 2015), and `hbsae` (Boonstra 2012) packages that use unit-level and area-level models such as the Empirical Best Linear Unbiased Prediction (EBLUP) estimation strategy and the hierarchical Bayesian estimation strategy. Rao (2003) discusses the benefits of the EBLUP for balancing potential bias of synthetic estimators against the instability of a direct estimator. White et al (2021) discusses the benefits of Small Area Estimation in a hierarchical Bayesian context, especially for forestry data. The module includes functional steps for checking, compiling, and formatting FIA plot data and ancillary spatial information for input to R packages, such as `JoSAE` (Breidenbach 2015), `sae` (Molina and Marhuenda 2015), or `hbsae` (Boonstra 2012) and translates integrated package output to `FIESTA` output format. 

Functions in `FIESTA` used for fitting Small Area Estimators include the `modSAarea` function for area estimates and `modSAtree` for tree estimates. The `modSApop` function is used to get population data needed for small area estimation. Below is a description and table of contents for the sections related to these functions:

FUNCTION  | DESCRIPTION
-------------- | ---------------------------------------------------------------
[modSApop](#modSApop) | Creates population data for small area estimation.
[modSAarea](#modSAarea) | Produces area level estimates through small area estimation.
[modSAtree](#modSAtree) | Produces tree level estimates through small area estimation.

## Objective of tutorial

The main objective of this tutorial is to demonstrate how to use `FIESTA` for generating estimates using estimators from the `JoSAE`, `sae`, and `hbsae` R packages. The following examples are for generating estimates and estimated variances using standard FIA Evaluation data from FIA's National database, with custom Estimation unit and Stratification information. The examples use data from three inventory years of field measurements in the state of Wyoming, from FIADB_1.7.2.00, last updated June 20, 2018, downloaded on June 25, 2018 and stored as internal data objects in FIESTA.


## Example data - Wyoming (WY), Inventory Years 2011-2012

Data Frame | Description
-----------| --------------------------------------------------------------------------------
WYplt      | WY plot-level data
WYcond     | WY condition-level data
WYtree     | WY tree-level data

External data            | Description
-------------------------| ------------------------------------------------------------------
WYbighorn_adminbnd.shp   | Polygon shapefile of WY Bighorn National Forest Administrative boundary*
WYbighorn_districtbnd.shp| Polygon shapefile of WY Bighorn National Forest District boundaries**
WYbighorn_forest_nonforest_250m.tif| GeoTIFF raster of predicted forest/nonforest (1/0) for stratification*** 
WYbighorn_dem_250m.img   | Erdas Imagine raster of elevation change, in meters****

*USDA Forest Service, Automated Lands Program (ALP). 2018. S_USA.AdministrativeForest (\url{http://data.fs.usda.gov/geodata/edw}). Description: An area encompassing all the National Forest System lands administered by an administrative unit. The area encompasses private lands, other governmental agency lands, and may contain National Forest System lands within the proclaimed boundaries of another administrative unit. All National Forest System lands fall within one and only one Administrative Forest Area.

**USDA Forest Service, Automated Lands Program (ALP). 2018. S_USA.RangerDistrict (http://data.fs.usda.gov/geodata/edw). Description: A depiction of the boundary that encompasses a Ranger District.

***Based on MODIS-based classified map resampled from 250m to 500m resolution and reclassified from 3 to 2 classes: 1:forest; 2:nonforest. Projected in Albers Conical Equal Area, Datum NAD27 (Ruefenacht et al. 2008). Clipped to extent of WYbighorn_adminbnd.shp.

****USGS National Elevation Dataset (NED), resampled from 30m resolution to 250m. Projected in Albers Conical Equal Area, Datum NAD27 (U.S. Geological Survey 2017). Clipped to boundary of WYbighorn_adminbnd.shp.


### Set up 

First, you'll need to load the `FIESTA` library:

```{r, warning = F, message = F}
library(FIESTA)
```

Next, you'll need to set up an "outfolder". This is just a file path to a folder where you'd like `FIESTA` to send your data output. For our purposes in this vignette, we have saved our outfolder file path as the `outfolder` object in a temporary directory. We also set a few default options preferred for this vignette.

```{r}
## Set options
options(scipen = 6)
options(stringsAsFactors = FALSE)

## Set outfolder
outfolder <- tempdir()
```


### Get data for examples

Now that we've loaded `FIESTA` and setup our outfolder, we can retrieve the data needed to run the examples. First, we point to some external data and predictor layers stored in `FIESTA` and derive new predictor layers using the `raster` package. 

```{r}
# File names for external spatial data
FIAplotsfn <- system.file("extdata", "FIApublic_M331B.db", package="FIESTA")
FIAplotsfn <- "C:/_tsf/_GitHub/FIESTA/inst/extdata/FIA_data/FIApublic_M331B.db"

WYbhfn <- system.file("extdata", "sp_data/WYbighorn_adminbnd.shp", package="FIESTA")
WYbhdistfn <- system.file("extdata", "sp_data/WYbighorn_districtbnd.shp", package="FIESTA")
WYbhdist.att <- "DISTRICTNA"

fornffn <- system.file("extdata", "sp_data/WYbighorn_forest_nonforest_250m.tif", package="FIESTA")
demfn <- system.file("extdata", "sp_data/WYbighorn_dem_250m.img", package="FIESTA")

# # Derive new layers from dem
# aspfn <- file.path(outfolder, "WYbh_asp.img")
# gdalUtils::gdaldem("aspect", input_dem=demfn, output=aspfn, of="GTiff")
# 
# slpfn <- file.path(outfolder, "WYbh_slp.img")
# gdalUtils::gdaldem("slope", input_dem=demfn, output=slpfn, of="GTiff")


# Derive new predictor layers from dem
library(raster)
dem <- raster(demfn)
slp <- terrain(dem,
               opt = "slope",
               unit = "degrees",
               filename = paste0(outfolder, "/WYbh_slp.img"), 
               overwrite = TRUE)
asp <- terrain(dem,
               opt = "aspect",
               unit = "degrees", 
               filename = paste0(outfolder, "/WYbh_asp.img"),
               overwrite = TRUE)
```

Next, we set up our small area domains with `FIESTA::spGetSAdoms`. For more information on how to use this function, please see the `sp` vignette included with `FIESTA` (link). 

```{r}
smallbnd <- WYbhdistfn
smallbnd.domain <- "DISTRICTNA"
```

Next, we can get our FIA plot data and set up our auxiliary data. We can get our FIA plot data with the `spGetPlots` function from `FIESTA`, which accesses data through [FIA's DataMart](https://apps.fs.usda.gov/fia/datamart/datamart.html). Here, data are downloaded for all states intersecting boundary, then subset to boundary. 

```{r}
SApltdat <- spGetPlots(bnd = WYbhdistfn,
                     xy_datsource = "obj",
                     xy = WYplt,
                     xy.uniqueid = "CN",
                     xvar = "LON_PUBLIC",
                     yvar = "LAT_PUBLIC",
                     xy.crs = 4269, 
                     datsource = "obj",
                     istree = TRUE,
                     isseed = TRUE,
                     plot_layer = WYplt,
                     cond_layer = WYcond,
                     tree_layer = WYtree,
                     seed_layer = WYseed,
                     invyrs = 2011:2013,
                     showsteps = TRUE,
                     savexy=TRUE,
                     savedata_opts = savedata_options(outfolder = outfolder))
```

```{r}
str(SApltdat, max.level = 1)
```

Finally, we must generate the dataset with predictors for small area estimation. We can do this with the `spGetAuxiliary` function from `FIESTA`. Again, see the `sp` vignette for further information on this function. 

```{r, results='hide'}
rastlst.cont <- c(dem, slp, asp)
rastlst.cont.name <- c("dem", "slp", "asp")
rastlst.cat <- fornffn
rastlst.cat.name <- "fornf"

unit_layer <- WYbhdistfn
unitvar <- "DISTRICTNA"

auxdat <- spGetAuxiliary(
  xyplt = SApltdat$spxy,
  uniqueid = "CN",
  unit_layer = unit_layer,
  unitvar = "DISTRICTNA",
  rastlst.cont = rastlst.cont,
  rastlst.cont.name = rastlst.cont.name,
  rastlst.cont.stat = "mean",
  rastlst.cont.NODATA = 0,
  rastlst.cat = rastlst.cat,
  rastlst.cat.name = rastlst.cat.name,
  asptransform = TRUE,
  rast.asp = asp,
  keepNA = FALSE,
  showext = FALSE,
  savedata = FALSE
)
names(auxdat)
```

```{r}
str(auxdat, max.level = 1)
```


## Examples

### `modSApop`

#### Example 1: Creating our population dataset with `modMApop`

We can create our population data for model-assisted estimation. To do so, we use the `modSApop` function in `FIESTA`. We must assign our plot data with the `pltdat` argument, the auxiliary dataset with the `auxdat` argument, and set information for our small areas with the `smallbnd` and `smallbnd.domain` arguments. The `spGetPlots` and `spGetAuxiliary` functions have done much of the hard work for us so far, so we can just run a simple call to `modSApop`:

```{r}
SApopdat <- modSApop(pltdat = SApltdat, 
                     auxdat = auxdat,
                     smallbnd = WYbhdistfn,
		                 smallbnd.domain = smallbnd.domain)
```

Note that the `modSApop` function returns a list with lots of information and data for us to use. For a quick look at what this list includes we can use the `str` function: 

```{r}
str(SApopdat, max.level = 1)
```

Now that we've created our population dataset, we can move on to estimation. 

### `modSAarea`

#### Example 2: Area of forest land, unit-level EBLUP

First, we can set up our predictors as a vector:

```{r}
all_preds <- c("slp", "dem", "asp_cos", "asp_sin", "fornf2")
```

Next, we fit the unit-level EBLUP with the `JoSAE` R package. 

```{r}
area1 <- modSAarea(
  SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
  prednames = all_preds, # est - charater vector of predictors to be used in the model
  SApackage = "JoSAE", # est - character string of the R package to do the estimation
  SAmethod = "unit" # est - method of small area estimation. Either "unit" or "area"
  )
```

The `modSAarea` function outputs a list, and we can see our estimates and estimation method.

```{r}
str(area1, max.level = 1)
area1$est
area1$raw$SAmethod
```

#### Example 3: Area of forest land, area-level EBLUP

In this example, we fit the same model as above, however now we can fit it at the area level. 

```{r}
area2 <- modSAarea(
  SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
  prednames = all_preds, # est - charater vector of predictors to be used in the model
  SApackage = "JoSAE", # est - character string of the R package to do the estimation
  SAmethod = "area" # est - method of small area estimation. Either "unit" or "area"
  )
```

We again can see our estimates, however with this data the area level model from `JoSAE` returned `NA`, so `FIESTA` will return the direct estimate. This can be seen by looking at the `SAmethod` object in the `raw` list. 

```{r}
area2$est
area2$raw$SAmethod
```

Since `FIESTA` will attempt fit all models when running `modSAarea`, we can look at all the different modeling approaches and their estimates with the `multest` object.

```{r}
area2$multest
```

Notably, area-level models return NAs with this data due to tiny values for within-area variation. So, for the rest of this vignette, we will focus on unit-level models. 

#### Example 4: Area of forest land, hierarchical Bayesian models

`FIESTA` also supports the use of hierarchical Bayesian (HB) models through the `hbsae` package as an alternative to EBLUPs. To fit these models, we can just change the `SApackage` argument.

```{r}
area3 <- modSAarea(
  SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
  prednames = all_preds, # est - charater vector of predictors to be used in the model
  SApackage = "hbsae", # est - character string of the R package to do the estimation
  SAmethod = "unit" # est - method of small area estimation. Either "unit" or "area"
  )
```

We can again check our estimates and small area method. 

```{r}
area3$est
area3$raw$SAmethod
```

#### Example 5: Ara of forest land, hierarchical Bayesian models, changing prior distribution

Notably, we can also set priors on the ratio of between and within area variation with `hbsae`. By default, `FIESTA` uses a weakly informative half-Cauchy prior on this parameter as suggested by White et al (2021), but in this example we will fit the same model as before, but with a flat prior. 

```{r}
area4 <- modSAarea(
  SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
  prednames = all_preds, # est - charater vector of predictors to be used in the model
  SApackage = "hbsae", # est - character string of the R package to do the estimation
  SAmethod = "unit", # est - method of small area estimation. Either "unit" or "area"
  prior = function(x) 1
  )
```

Let's check our results compared to Example 3 (same model with half-Cauchy prior)

```{r}
area3$est
area4$est
```

Due to rounding, we see the same result. However, the estimates are slightly different. 

#### Example 6: Area of forest land, with model variable selection

### `modSAtree`



```{r}
# test1 <- modSAarea(
#   SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
#   prednames = c("slp", "dem", "asp_cos", "asp_sin"),
#   SApackage = "hbsae",
#   SAmethod = "area",
#  # variable.select = FALSE,
#   totals = TRUE)
# 
# test2 <- modSAarea(
#   SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
#   prednames = c("slp", "dem", "asp_cos", "asp_sin"),
#   SApackage = "hbsae",
#   SAmethod = "unit",
#   variable.select = FALSE,
#   totals = TRUE)

# 
# area1$est
# area1$raw$SAmethod
```

#### Example 5:

```{r}
# area1 <- modSAarea(
#   SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
#   prednames = c("slp", "dem", "asp_cos", "asp_sin"),
#   SApackage = "hbsae",
#   SAmethod = "unit",
#   variable.select = FALSE
#   )
# 
# area1$est
# area1$raw$SAmethod
```







#### <a name="modSAtree"/>modSAtree()
```{r}


tree1 <- modSAtree(
  SApopdatlst = SApopdat, # pop - population calculations for WY, post-stratification
  prednames = c("slp", "dem", "asp_cos", "asp_sin"),
  SApackage = "JoSAE",
  SAmethod = "unit",
  estvar = "BA"
)

tree1$raw$SAmethod

```



## References

Breidenbach J. 2018. JoSAE: Unit-Level and Area-Level Small Area Estimation.

Molina I, Marhuenda Y. 2015. sae: An R Package for Small Area Estimation. The R Journal, 7(1), 81–98. https://journal.r-project.org/archive/2015/RJ-2015-007/RJ-2015-007.pdf.

Rao, J.N.K. 2003. Small Area Estimation. Wiley, Hoboken, New Jersey.

