

# Using auxiliary data in FIESTA can help enhance your result. One resource for auxiliary data is EarthData Search: a NASA repository for Earth Science data.

## Follow this tutorial and learn how to implement auxiliary data from EarthData Search (https://search.earthdata.nasa.gov/search)

# ############################################################

# SET UP

# ############################################################


## This section will include the following

* Set up R packages
* Set up user settings and files
* Set up download file

## Set up R packages
```{r setup, include = F}
library(rsi)
library("FIESTA")
```


```{r setup, include = F}
spectral_indices()

aoi <- sf::st_point(c(-107.34041007184679, 44.48965583058798))
aoi <- sf::st_set_crs(sf::st_sfc(aoi), 4326)
aoi <- sf::st_buffer(sf::st_transform(aoi, 5070), 10)

landsat_image_red <- get_stac_data(
  aoi,
  start_date = "2022-06-01",
  end_date = "2022-06-30",
  pixel_x_size = 30,
  pixel_y_size = 30,
  asset_names = c("red"),
  stac_source = "https://planetarycomputer.microsoft.com/api/stac/v1/",
  collection = "landsat-c2-l2",
  mask_band = "qa_pixel",
  mask_function = landsat_mask_function,
  output_filename = tempfile(fileext = ".tif"),
  item_filter_function = landsat_platform_filter,
  platforms = c("landsat-9", "landsat-8")
)

landsat_image_green <- get_stac_data(
  aoi,
  start_date = "2022-06-01",
  end_date = "2022-06-30",
  pixel_x_size = 30,
  pixel_y_size = 30,
  asset_names = c("green"),
  stac_source = "https://planetarycomputer.microsoft.com/api/stac/v1/",
  collection = "landsat-c2-l2",
  mask_band = "qa_pixel",
  mask_function = landsat_mask_function,
  output_filename = tempfile(fileext = ".tif"),
  item_filter_function = landsat_platform_filter,
  platforms = c("landsat-9", "landsat-8")
)

landsat_image_blue <- get_stac_data(
  aoi,
  start_date = "2022-06-01",
  end_date = "2022-06-30",
  pixel_x_size = 30,
  pixel_y_size = 30,
  asset_names = c("blue"),
  stac_source = "https://planetarycomputer.microsoft.com/api/stac/v1/",
  collection = "landsat-c2-l2",
  mask_band = "qa_pixel",
  mask_function = landsat_mask_function,
  output_filename = tempfile(fileext = ".tif"),
  item_filter_function = landsat_platform_filter,
  platforms = c("landsat-9", "landsat-8")
)

terra::plot(terra::rast(landsat_image))

```


# ############################################################

# USE DATA IN FIESTA

# ############################################################

## Translate downloaded data to Auxiliary data
```{r setup, include = F}

# Set up data necessary for spGetAuxiliary function
WYbhfn <- system.file("extdata",
                      "sp_data/WYbighorn_adminbnd.shp",
                      package = "FIESTA")

# Import downloaded data as raster file
library(terra)
data_red <- rast(landsat_image_red)
data_green <- rast(landsat_image_green)
data_blue <- rast(landsat_image_blue)

# Get Wyoming Plots
WYspplt <- spMakeSpatialPoints(xyplt = WYplt,
                               xy.uniqueid = "CN",
                               xvar = "LON_PUBLIC",
                               yvar = "LAT_PUBLIC",
                               xy.crs = 4269)

# Run sp Get Auxiliary
rastlst.cont <- c(data_red, data_green, data_blue)
rastlst.cont.name <- c("red", "green", "blue")

auxiliaryData <- spGetAuxiliary(xyplt = WYspplt,
                                uniqueid = "CN",
                                unit_layer = WYbhfn,
                                unitvar = NULL,
                                rastlst.cont = rastlst.cont,
                                rastlst.cont.name = rastlst.cont.name,
                                rastlst.cont.stat = "mean",
                                keepNA = FALSE,
                                showext = FALSE,
                                savedata = FALSE)

```



## Use auxiliary data in Small Area population
```{r setup, include = F}
# Setting up necessary data for modSApop function
WYbhdistfn <- system.file("extdata", "sp_data/WYbighorn_districtbnd.shp", package="FIESTA")

smallbnd <- WYbhdistfn
smallbnd.domain <- "DISTRICTNA"

SApltdat <- spGetPlots(bnd = WYbhdistfn,
                       xy_datsource = "obj",
                       xy = WYplt,
                       xy_opts = list(xy.uniqueid = "CN", xvar = "LON_PUBLIC", 
                                      yvar = "LAT_PUBLIC", xy.crs = 4269),
                       datsource = "obj",
                       istree = TRUE,
                       isseed = TRUE,
                       dbTabs = list(plot_layer = WYplt, cond_layer = WYcond,
                                     tree_layer = WYtree, seed_layer = WYseed),
                       eval = "custom",
                       eval_opts = list(invyrs = 2011:2013),
                       showsteps = TRUE,
                       returnxy = TRUE)

# Use Auxiliary Data in modSApop
SApopdat <- modSApop(pltdat = SApltdat, auxdat = auxiliaryData,
                     smallbnd = WYbhdistfn, smallbnd.domain = smallbnd.domain)
```






