# Set up r
```{r setup, include = F}
library(httr)
library("FIESTA")
```


# Set up user settings and download paths

## Add these files to the auxdata_example directory and add your credentials

### Filename: .netrc
### Contents: machine urs.earthdata.nasa.gov login uid_goes_here password password_goes_here

### Filename: .urs_cookies
### Contents: empty

```{r setup, include = F}
netrc_path <- "./auxdata_example/.netrc"
cookie_path <- "./auxdata_example/.urs_cookies"
download_folder <- "./auxdata_example/downloads"
```


# Set data and download location

## This is where you can change the url of the data you are downloading as well as the downloaded filename
```{r setup, include = F}

data_url <- "https://data.lpdaac.earthdatacloud.nasa.gov/lp-prod-protected/HLSS30.020/HLS.S30.T13TCK.2023364T180741.v2.0/HLS.S30.T13TCK.2023364T180741.v2.0.B03.tif"
filename <- "landsat.tif"
```


# Download the Data
```{r setup, include = F}
# Set Configuration for authentication
set_config(config(followlocation = 1,
                  netrc = 1,
                  netrc_file = netrc_path,
                  cookie = cookie_path,
                  cookiefile = cookie_path,
                  cookiejar = cookie_path))

downloaded_file_path <- paste(download_folder, "/", filename, sep = "")

httr::GET(url = data_url, write_disk(downloaded_file_path, overwrite = TRUE))
```


# Translate downloaded data to Auxiliary data
```{r setup, include = F}
WYbhfn <- system.file("extdata",
                      "sp_data/WYbighorn_adminbnd.shp",
                      package = "FIESTA")

library(terra)
data <- rast(downloaded_file_path)
data_name <- "Input_Data"

# Get Wyoming Plots
WYspplt <- spMakeSpatialPoints(xyplt = WYplt,
                                xy.uniqueid = "CN",
                                xvar = "LON_PUBLIC",
                                yvar = "LAT_PUBLIC",
                                xy.crs = 4269)

# Run sp Get Auxiliary
rastlst.cont <- c(data)
rastlst.cont.name <- c(data_name)

auxiliaryData <- spGetAuxiliary(xyplt = WYspplt,
                              uniqueid = "CN",
                              unit_layer = WYbhfn,
                              unitvar = NULL,
                              rastlst.cont = rastlst.cont,
                              rastlst.cont.name = rastlst.cont.name,
                              rastlst.cont.stat = "mean",
                              keepNA = FALSE,
                              showext = FALSE,
                              savedata = FALSE)

```


# Use auxiliary data in Small Area population
```{r setup, include = F}

WYbhdistfn <- system.file("extdata", "sp_data/WYbighorn_districtbnd.shp", package="FIESTA")

smallbnd <- WYbhdistfn
smallbnd.domain <- "DISTRICTNA"

SApltdat <- spGetPlots(bnd = WYbhdistfn,
                       xy_datsource = "obj",
                       xy = WYplt,
                       xy_opts = list(xy.uniqueid = "CN", xvar = "LON_PUBLIC", 
                                    yvar = "LAT_PUBLIC", xy.crs = 4269),
                       datsource = "obj",
                       istree = TRUE,
                       isseed = TRUE,
                       dbTabs = list(plot_layer = WYplt, cond_layer = WYcond,
                                   tree_layer = WYtree, seed_layer = WYseed),
                       eval = "custom",
                       eval_opts = list(invyrs = 2011:2013),
                       showsteps = TRUE,
                       returnxy = TRUE)

SApopdat <- modSApop(pltdat = SApltdat,
                     auxdat = auxiliaryData,
                     smallbnd = WYbhdistfn,
		             smallbnd.domain = smallbnd.domain)

```