

# Using auxiliary data in FIESTA can help enhance your result. One resource for auxiliary data is EarthData Search: a NASA repository for Earth Science data.

## Follow this tutorial and learn how to implement auxiliary data from EarthData Search (https://search.earthdata.nasa.gov/search)

# ############################################################

# SET UP

# ############################################################


## This section will include the following

* Set up R packages
* Set up user settings and files
* Set up download file

## Set up R packages
```{r setup, include = F}
library(httr)
library("FIESTA")
```


## Set up user settings and files

### To use EarthData Search, one must have an account. Go to EarthData Search and create an account and note your username and password

```{r setup, include = F}
# Create a '.netrc' file at the below location (This file is untracked in git repositories)
# The contents of the .netrc file should follow this template:
# machine urs.earthdata.nasa.gov login YOUR_USER_NAME password YOUR_PASSWORD

netrc_path <- "./auxdata_example/.netrc"

# Next create a .urs_cookies file at the below location and leave it blank (This file is untracked in git repositories)
cookie_path <- "./auxdata_example/.urs_cookies"


# For more information on setting up an .netrc file go to: https://wiki.earthdata.nasa.gov/display/EL/How+To+Access+Data+With+cURL+And+Wget
# For more information on using EarthData Search with R go to: https://urs.earthdata.nasa.gov/documentation/for_users/data_access/r
```


## Set up download file

### This is where you can change the url of the data you are downloading as well as the downloaded filename
```{r setup, include = F}

# Choose a folder where you would like the data to download
download_folder <- tempdir()

# Paste in the download link of the data (This is the download link from EarthData Search *Right click download and save url*)
data_url <- "https://data.lpdaac.earthdatacloud.nasa.gov/lp-prod-protected/HLSS30.020/HLS.S30.T13TCK.2023364T180741.v2.0/HLS.S30.T13TCK.2023364T180741.v2.0.B03.tif" # Harmonized Landsat Sentinel Product Band 3

# Create a filename for the downloaded data
filename <- "HLS_B3.tif"
```


# ############################################################

# DOWNLOAD THE DATA

# ############################################################

```{r setup, include = F}
# Set Configuration for authentication
set_config(config(followlocation = 1,
                  netrc = 1,
                  netrc_file = netrc_path,
                  cookie = cookie_path,
                  cookiefile = cookie_path,
                  cookiejar = cookie_path))

# Create Full File Path
downloaded_file_path <- paste(download_folder, "/", filename, sep = "")

# Download File (May take time)
httr::GET(url = data_url, write_disk(downloaded_file_path, overwrite = TRUE))
```


# ############################################################

# USE DATA IN FIESTA

# ############################################################

## Translate downloaded data to Auxiliary data
```{r setup, include = F}

# Set up data necessary for spGetAuxiliary function
WYbhfn <- system.file("extdata",
                      "sp_data/WYbighorn_adminbnd.shp",
                      package = "FIESTA")

# Import downloaded data as raster file
library(terra)
data <- rast(downloaded_file_path)
data_name <- "Input_Data"

# Get Wyoming Plots
WYspplt <- spMakeSpatialPoints(xyplt = WYplt,
                                xy.uniqueid = "CN",
                                xvar = "LON_PUBLIC",
                                yvar = "LAT_PUBLIC",
                                xy.crs = 4269)

# Run sp Get Auxiliary
rastlst.cont <- c(data)
rastlst.cont.name <- c(data_name)

auxiliaryData <- spGetAuxiliary(xyplt = WYspplt,
                              uniqueid = "CN",
                              unit_layer = WYbhfn,
                              unitvar = NULL,
                              rastlst.cont = rastlst.cont,
                              rastlst.cont.name = rastlst.cont.name,
                              rastlst.cont.stat = "mean",
                              keepNA = FALSE,
                              showext = FALSE,
                              savedata = FALSE)

```


## Use auxiliary data in Small Area population
```{r setup, include = F}
# Setting up necessary data for modSApop function
WYbhdistfn <- system.file("extdata", "sp_data/WYbighorn_districtbnd.shp", package="FIESTA")

smallbnd <- WYbhdistfn
smallbnd.domain <- "DISTRICTNA"

SApltdat <- spGetPlots(bnd = WYbhdistfn,
                       xy_datsource = "obj",
                       xy = WYplt,
                       xy_opts = list(xy.uniqueid = "CN", xvar = "LON_PUBLIC", 
                                    yvar = "LAT_PUBLIC", xy.crs = 4269),
                       datsource = "obj",
                       istree = TRUE,
                       isseed = TRUE,
                       dbTabs = list(plot_layer = WYplt, cond_layer = WYcond,
                                   tree_layer = WYtree, seed_layer = WYseed),
                       eval = "custom",
                       eval_opts = list(invyrs = 2011:2013),
                       showsteps = TRUE,
                       returnxy = TRUE)

# Use Auxiliary Data in modSApop
SApopdat <- modSApop(pltdat = SApltdat, auxdat = auxiliaryData,
                     smallbnd = WYbhdistfn, smallbnd.domain = smallbnd.domain)
```