---
output:
  word_document:
    reference_docx: anGBtemplate.docx
params:
  SApopdatlst: ''
  SApackage: ''
  SAmethod: ''
  AOInm: ''
  pcfilter: ''
  fortypgrpcd: ''
  title.ref: ''
  totals: ''

---


```{r Params, echo=FALSE, warning=FALSE, message=FALSE}
 
library(data.table)
library(RColorBrewer)

SApopdatlst <- params$SApopdatlst
AOInm <- params$AOInm
pcfilter <- params$pcfilter
SApackage <- params$SApackage
SAmethod <- params$SAmethod
fortypgrpcd <- params$fortypgrpcd
title.ref <- params$title.ref
#domain <- params$domain
totals <- params$totals

SApopdatlst = SApopdatlst 
SApackage = SApackage
SAmethod = SAmethod 
AOInm = AOInm
title.ref = title.ref
totals = FALSE
outfolder = outfolder

```


```{r Data_Inputs, echo=FALSE, warning=FALSE, message=FALSE}

digits <- c(0,0,4,4,2,0)
na.fill <- "DIR"
modelselect <- TRUE

# AOInm <- "ecoM332"
# load("E:/workspace/FIESTA/FIESTA_SA/testfolder4/M332/SApopdatlst.rda")
# pcfilter <- NULL
# title.ref <- NULL
# fortypgrpcd <- 200
# estnm="JFH"
# senm="JFH.se"
# domain="Incident_Name"
 

if (is.null(title.ref)) {
  title.ref <- AOInm
}

eval <- TRUE
if (is.null(SApopdatlst[[1]]$smallbnd)) {
  eval <- FALSE
}

smallbnd.domain <- SApopdatlst[[1]]$smallbnd.domain
if (is.null(smallbnd.domain)) {
  smallbnd.domain <- "DOMAIN"
}


## pcfilter
for (i in 1:length(SApopdatlst)) {
  if (!is.null(pcfilter) && 'STATECD' %in% names(SApopdatlst$pltcondx)) {
    pltcondf <- datFilter(SApopdat$pltcondx, xfilter=pcfilter)$xf
    states <- pcheck.states(unique(pltcondf$STATECD))
    invyrs <- unique(pltcondf$INVYR)
  }
}

## Get all states and inventory years in SApopdatlst 
states <- {}
invyrs <- {}
for (i in 1:length(SApopdatlst)) {
  if (!is.null(pcfilter) && 'STATECD' %in% names(SApopdatlst[[i]]$pltcondx)) {
    pltcondf <- datFilter(SApopdatlst[[i]]$pltcondx, xfilter=pcfilter)$xf
    states <- sort(c(states, pcheck.states(unique(pltcondf$STATECD))))
    invyrs <- sort(unique(c(invyrs, pltcondf$INVYR)))
  } else if (!is.null(SApopdatlst[[i]]$states)) {
    states <- sort(unique(c(states, SApopdatlst[[i]]$states)))
    invyrs <- sort(unique(c(invyrs, unlist(SApopdatlst[[i]]$invyrs))))
  }
}


if (is.null(pcfilter)) {
  title.statesx=title.states <- NULL
  title.filter <- NULL
} else {
  if (length(states) < 3) {
    title.statesx <- paste(states, collapse=" and ")
    title.states <- paste0(", ", paste(states, collapse=" and "))
  } else {
    title.statesx=title.states <- NULL 
  }
  title.filter <- AOInm
}
invyrs2 <- paste0(min(invyrs), "-", max(invyrs))


if (!eval) {
  intro_text1 <- "" 
                      
  intro_text2 <-
    paste0("Estimates are calculated using the Forest Inventory ESTimation and Analysis (FIESTA) R package (Frescino et al. 2015) and are based on area-level small area estimators from the R packages. Data for this report were extracted from the Forest Inventory and Analysis (FIA) database (Burrill et al. 2018).")
}

```


---
title: "Forest Resource Report:  \n`r AOInm``r title.states`,  \n`r invyrs2`"
---

-----  

&nbsp;  

Rocky Mountain Research Station, Forest Inventory and Analysis  | `r format(Sys.time(), "%B %Y")`

&nbsp;  

```{r smallbnd, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7.5, eval=eval}

smallbnd <- {}
for (i in 1:length(SApopdatlst)) {
  smallbnd <- rbind(smallbnd, SApopdatlst[[i]]$smallbnd)
  smallbnd.unique <- SApopdatlst[[i]]$smallbnd.unique
}  

plot(sf::st_geometry(smallbnd), border="black", main="")

```
Figure 1. Small area of interest, `r title.ref``r title.states`, `r invyrs2`.

&nbsp;  


----- 


# Summary

&nbsp;  

```{r Summary_Info, echo=FALSE, include=FALSE}
 

dunitlut <- {}
dunitarea <- {}
for (i in 1:length(SApopdatlst)) {
  if (!"AOI" %in% names(SApopdatlst[[i]]$dunitlut)) {
    SApopdatlst[[i]]$dunitarea$AOI <- 1
    SApopdatlst[[i]]$dunitlut$AOI <- 1
    SApopdatlst[[i]]$condx$AOI <- 1
  }
    
  dunitlut <- rbind(dunitlut, SApopdatlst[[i]]$dunitlut)
  dunitarea <- rbind(dunitarea, SApopdatlst[[i]]$dunitarea)
  areavar <- SApopdatlst[[i]]$areavar
}


dominfo <- merge(dunitlut[, c("DOMAIN", "AOI", "n.total")], 
                  dunitarea, by=c("DOMAIN", "AOI"))
nbrdoms <- nrow(dominfo)
totarea <- sum(dominfo[dominfo$AOI == 1, areavar, with=FALSE])

#knitr::kable(dominfo, "simple", align=c("lrr"),
#              digits=digits, format.args=list(big.mark = ","))


intro_text1 <- paste0("The total area of ", AOInm, " is ", format(round(totarea), big.mark=","), 
        " acres with ", nbrdoms, " unique small areas. The area intersects the following states: ",
        toString(states), ".") 
                      

intro_text2 <-
paste0("Estimates are calculated using the Forest Inventory ESTimation and Analysis (FIESTA) R package (Frescino et al. 2015) and are based on area-level small area estimators from the R packages. Data for this report were extracted from the Forest Inventory and Analysis (FIA) database for years ", invyrs2, " (Burrill et al. 2018).")

ndoms <- nrow(dunitlut)
figwidth <- ifelse(ndoms < 4, "25%", ifelse(ndoms < 8, "50%", 
                   ifelse(ndoms < 12, "75%", "100%")))

```


`r intro_text1`

&nbsp;  

`r intro_text2`

&nbsp;   

\newpage

&nbsp;  

-----  

# EST1 - Area of forest land by `r tolower(smallbnd.domain)`.

-----  

&nbsp;  

Table 1. Area of forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`.

&nbsp;  


```{r echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
SAareadat1 <- modSAarea(SApopdatlst = SApopdatlst, 
                        SApackage = SApackage, 
                        SAmethod = SAmethod, 
                        landarea = "FOREST", 
                        modelselect = modelselect,
                        na.fill = na.fill,
                        returntitle = TRUE, 
                        table_opts = list(totals=totals),
                        title_opts = list(title.ref=title.ref))
#est1 <- SAareadat1$est
#multest1 <- SAareadat1$multest[SAareadat1$multest$AOI == 1, ]
est1 <- SAareadat1$raw$dunit_totest
titlelst1 <- SAareadat1$titlelst
SAmethod_used1 <- SAareadat1$raw$SAmethod

#multest1$pse <- multest1[[senm]] / multest1[[estnm]] * 100
est1 <- est1[, c("DOMAIN", "NBRPLT.gt0", "est", "est.se", "pse", "AREAUSED")]
names(est1)[names(est1) == "DOMAIN"] <- smallbnd.domain

```  

```{r est1}
knitr::kable(est1, "simple", align=c("lrrrrr"), digits=digits,
             format.args=list(big.mark = ","))
```

&nbsp;  


&nbsp; 


```{r EST1_Barplot, echo=FALSE, warning=FALSE, message=FALSE, out.width=figwidth}

est1tot <- est1
if (totals) {
  ylab <- "Acres"
} else {
  ylab <- "Per Acre"
}

if (length(est1tot$est) > 1) {
#horiz <- ifelse(nrow(est1tot) > 10, TRUE, FALSE)
  horiz <- FALSE
  datBarplot(x=est1tot, xvar=smallbnd.domain, yvar="est", horiz=horiz,
           errbars=TRUE, sevar="est.se", toplabelvar="NBRPLT.gt0", ylab=ylab)
}
```
<!-- Figure 1. Area of forest land by, `r tolower(smallbnd.domain)`, `r AOInm`, `r title.states`, `r invyrs2`. -->


&nbsp;    


```{r est1_map, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7.5, eval=eval}


## Merge estimate to smallbnd
smallbnd1 <- merge(smallbnd, est1, by=smallbnd.domain)

pal <- palette(brewer.pal(n=5, name = "Greens"))

ptest <- tryCatch(
plot(smallbnd1["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=5, 
 	pal=palette(brewer.pal(n=5, name="Greens")),
 	border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })

if (is.null(ptest)) {
  n <- nrow(smallbnd1["est"])
  
  pal <- palette(brewer.pal(n=n, name = "Greens"))

  ptest <- tryCatch(
    plot(smallbnd1["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	    graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=4, 
 	    pal=palette(brewer.pal(n=n, name="Greens")),
 	    border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })
}


# plot(smallbnd1[estnm], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
# 	graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=5, 
# 	border="black", lwd=.5, main="")


```
<!-- Figure 1. Area of forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`. -->


\newpage    

&nbsp;  


-----  

# EST2 - Net cubic-foot volume of live trees on forest land by `r tolower(smallbnd.domain)`.

-----  


&nbsp; 


Table 2. Net cubic-foot volume of live trees on forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`.

&nbsp;    

```{r echo=FALSE, warning=FALSE, message=FALSE, include = FALSE}
 
SAareadat2 <- modSAtree(SApopdatlst = SApopdatlst, 
                        SApackage = SApackage, 
                        SAmethod = SAmethod, 
                        landarea = "FOREST", 
                        estvar = "VOLCFNET", 
                        estvar.filter = "STATUSCD == 1", 
                        na.fill = na.fill,
                        returntitle = TRUE, 
                        table_opts = list(totals=totals),
                        title_opts = list(title.ref=title.ref))
#est2 <- SAareadat2$est
#multest2 <- SAareadat2$multest[SAareadat2$multest$AOI == 1, ]
est2 <- SAareadat2$raw$dunit_totest
titlelst2 <- SAareadat2$titlelst
SAmethod_used2 <- SAareadat2$raw$SAmethod

#multest2$pse <- multest2[[senm]] / multest2[[estnm]] * 100
est2 <- est2[, c("DOMAIN", "NBRPLT.gt0", "est", "est.se", "pse", "AREAUSED")]
names(est2)[names(est2) == "DOMAIN"] <- smallbnd.domain

```  

```{r est2}
knitr::kable(est2, "simple", align=c("lrrrrr"), digits=digits,
             format.args=list(big.mark = ","))
```


&nbsp;    


&nbsp;    


```{r EST2_Barplot, echo=FALSE, warning=FALSE, message=FALSE, out.width=figwidth}

est2tot <- est2
if (totals) {
  divideby <- "million"
  ylab <- "Acres"
} else {
  divideby <- NULL
  ylab <- "Per Acre"
}

if (length(est2tot$est) > 1) {

  horiz <- FALSE
  datBarplot(x=est2tot, xvar=smallbnd.domain, yvar="est", horiz=horiz,
           errbars=TRUE, sevar="est.se", toplabelvar="NBRPLT.gt0", 
           ylab=ylab, divideby=divideby)
}
```


&nbsp;  


```{r est2_map, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7.5, eval=eval}

## Merge estimate to smallbnd
smallbnd2 <- merge(smallbnd, est2, by=smallbnd.domain)

ptest <- tryCatch(
plot(smallbnd2["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=5, 
 	pal=brewer.pal(n=4, name="Reds"),
 	border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })

if (is.null(ptest)) {
  n <- nrow(smallbnd2["est"])

  ptest <- tryCatch(
    plot(smallbnd2["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	    graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=4, 
 	    pal=brewer.pal(n=n, name="Reds"),
 	    border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })
}

```
<!-- Figure 2. Net cubic-foot volume of live trees on forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`. -->


\newpage 

&nbsp;  

-----  

# EST3 - Square-feet basal area of live trees on forest land by `r tolower(smallbnd.domain)`.

-----  

Table 3. Square-feet basal area of live trees on forest land by `r tolower(smallbnd.domain)`, `r AOInm``r title.states`, `r invyrs2`.

&nbsp;    

```{r echo=FALSE, warning=FALSE, message=FALSE, include = FALSE}

SAareadat3 <- modSAtree(SApopdatlst = SApopdatlst, 
                       SApackage = SApackage, 
                       SAmethod = SAmethod, 
                       landarea = "FOREST", 
                       estvar = "BA", 
                       estvar.filter = "STATUSCD == 1", 
                       na.fill = na.fill, 
                       returntitle = TRUE, 
                       table_opts = list(totals=totals),
                       title_opts = list(title.ref = title.ref))
#multest3 <- SAareadat3$multest[SAareadat3$multest$AOI == 1, ]
est3 <- SAareadat3$raw$dunit_totest
titlelst3 <- SAareadat3$titlelst
SAmethod_used3 <- SAareadat3$raw$SAmethod

#multest3$pse <- multest3[[senm]] / multest3[[estnm]] * 100
est3 <- est3[, c("DOMAIN", "NBRPLT.gt0", "est", "est.se", "pse", "AREAUSED")]
names(est3)[names(est3) == "DOMAIN"] <- smallbnd.domain
```  


```{r est3}
knitr::kable(est3, "simple", align=c("lrrrr"), digits=digits,
             format.args=list(big.mark = ","))

```


&nbsp;  


&nbsp;   


```{r EST3_Barplot, echo=FALSE, warning=FALSE, message=FALSE, out.width=figwidth}


est3tot <- est3
if (totals) {
  divideby <- "million"
  ylab <- "Acres"
} else {
  divideby <- NULL
  ylab <- "Per acre"
}

if (length(est3tot$est) > 1) {

  datBarplot(x=est3tot, xvar=smallbnd.domain, yvar="est", 
           errbars=TRUE, sevar="est.se", toplabelvar="NBRPLT.gt0", 
           ylab=ylab, divideby=divideby)
}
```


&nbsp;    


```{r est3_map, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7.5, eval=eval}

smallbnd3 <- merge(smallbnd, est3, by=smallbnd.domain)

ptest <- tryCatch(
plot(smallbnd3["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=5, 
 	pal=brewer.pal(n=4, name="Blues"),
 	border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })

if (is.null(ptest)) {
  n <- nrow(smallbnd3["est"])

  ptest <- tryCatch(
    plot(smallbnd3["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	    graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=4, 
 	    pal=brewer.pal(n=n, name="Blues"),
 	    border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })
}

```
<!-- Figure 3. Square-feet basal area of live trees on forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`. -->


\newpage    

&nbsp;  


-----  

# EST4 - Net cubic-board foot volume of live trees on forest land by `r tolower(smallbnd.domain)`.

-----  

&nbsp; 

Table 4. Net board-foot volume of live trees on forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`.

&nbsp;    

```{r echo=FALSE, warning=FALSE, message=FALSE, include = FALSE}

SAareadat4 <- modSAtree(SApopdatlst = SApopdatlst, 
                       SApackage = SApackage, 
                       SAmethod = SAmethod, 
                       landarea = "FOREST", 
                       estvar = "VOLBFNET", 
                       estvar.filter = "STATUSCD == 1", 
                       na.fill = na.fill, 
                       returntitle = TRUE, 
                       table_opts = list(totals=totals),
                       title_opts = list(title.ref=title.ref))
#multest4 <- SAareadat4$multest[SAareadat4$multest$AOI == 1, ]
est4 <- SAareadat4$raw$dunit_totest
titlelst4 <- SAareadat4$titlelst
SAmethod_used4 <- SAareadat4$raw$SAmethod

#multest4$pse <- multest4[[senm]] / multest4[[estnm]] * 100
est4 <- est4[, c("DOMAIN", "NBRPLT.gt0", "est", "est.se", "pse", "AREAUSED")]
names(est4)[names(est4) == "DOMAIN"] <- smallbnd.domain

```  


```{r est4}
knitr::kable(est4, "simple", align=c("lrrrrr"), digits=digits,
             format.args=list(big.mark = ","))
```


&nbsp;    


&nbsp;    


```{r EST4_Barplot, echo=FALSE, warning=FALSE, message=FALSE, out.width=figwidth}

est4tot <- est4
if (totals) {
  divideby <- "million"
  ylab <- "Acres"
} else {
  divideby <- NULL
  ylab <- "Per acre"
}

if (length(est4tot$est) > 1) {

  datBarplot(x=est4tot, xvar=smallbnd.domain, yvar="est", 
           errbars=TRUE, sevar="est.se", toplabelvar="NBRPLT.gt0", 
           ylab=ylab, divideby=divideby)
}
```


&nbsp;    


```{r est4_map, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7.5, eval=eval}

## Merge estimate to smallbnd
smallbnd4 <- merge(smallbnd, est4, by=smallbnd.domain)

ptest <- tryCatch(
plot(smallbnd4["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3),
	graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=5,
	pal = brewer.pal(n=6, name="Oranges"),
	border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })

if (is.null(ptest)) {
  n <- nrow(smallbnd4["est"])

  ptest <- tryCatch(
    plot(smallbnd4["est"], key.pos=1, axes=TRUE, key.width=lcm(1.3), 
 	    graticule=TRUE, key.length=1.0, breaks="equal", nbreaks=4, 
 	    pal=brewer.pal(n=n, name="Oranges"),
 	    border="black", lwd=.5, main=""),
     	 	error=function(e) {
			return(NULL) })
}

```
<!-- Figure 4. Net board-foot volume of live trees on forest land by `r tolower(smallbnd.domain)`, `r title.ref``r title.states`, `r invyrs2`. -->


\newpage 



# References

&nbsp;  

Breidenbach, Johannes 2018. JoSAE: Unit-Level and Area-Level Small Area Estimation. R package version 0.3.0. https://CRAN.R-project.org/package=JoSAE.
  
&nbsp;

Burrill, E.A., Wilson, A.M., Turner, J.A., Pugh, S.A., Menlove, J., Christiansen, G., Conkling, B.L., Winnie, D., 2018. Forest Inventory and Analysis Database [WWW Document]. St Paul MN US Dep. Agric. For. Serv. North. Res. Stn. URL http://apps.fs.fed.us/fiadb-downloads/datamart.html (accessed 3.6.21).

&nbsp; 

Frescino, Tracey S.; Patterson, Paul L.; Moisen, Gretchen G.; Toney, Chris; Freeman, Elizabeth A. 2018. FIESTA: A Forest Inventory Estimation and Analysis R Package. USDA Forest Service, Rocky Mountain Research Station, 507 25th street, Ogden, UT, USA.