---
title: "Spatial Tools"
output: rmarkdown::html_vignette
description: >
  Explore the various spatial utility functions available within FIESTA.
vignette: >
  %\VignetteIndexEntry{Spatial Tools}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = F}
library(knitr)
knitr::opts_chunk$set(message = F, warning = F, eval=F)
```

```{r, include=FALSE}
# Sets up output folding
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold.", type)]])) return(res)
    
    paste0(
      "<details><summary>", type, "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot")
)
```

```{r, echo=-1}
data.table::setDTthreads(2)
```


## `FIESTA` Overview
The R package, `FIESTA` (Forest Inventory ESTimation and Analysis) is a research estimation tool for analysts that work with sample-based inventory data from the U.S. Department of Agriculture, Forest Service, Forest Inventory and Analysis (FIA) Program to accommodate: unique population boundaries, different evaluation time periods, customized stratification schemes, non-standard variance equations, integration of multi-scale remotely-sensed data and other ancillary information, and interaction with other modeling and estimation tools from CRAN R's library of packages. `FIESTA` contains a collection of functions that can access FIA databases, summarize and compile plot and spatial data, and generate estimates with associated sampling errors. 


Functions are organized by type or objective and are named with a corresponding prefix: 

**Core Functions**

* Database tools (`DB`) - functions for querying and extracting data from FIA's national database.
* Data tools (`dat`) - functions for summarizing and exploring FIA data.
* Spatial tools (`sp`) - functions for manipulating and summarizing spatial data.

**Estimation Modules**

* Green-Book (`GB`) - functions for FIA’s standard ‘Green-Book’ estimators.
* Photo-Based (`PB`) - functions for supplementary photo-based estimators.
* Small Area (`SA`) - functions for integration with available small area estimators (SAE).
* Model-Assisted (`MA`) - functions for integration with available Model-Assisted estimators.

**Analysis Tools**

* Analysis tools (`an`) - wrapper functions for stream-lining estimation processes.


## Overview of `FIESTA` spatial (`sp`) tools
`FIESTA`'s spatial tools allow for summarizing and manipulating spatial data for use in `FIESTA`'s estimation modules. 


FUNCTION  | DESCRIPTION
-------------- | ---------------------------------------------------------------
[spImportSpatial()](#spImportSpatial) | Imports a spatial layer to a sf object.
[spExportSpatial()](#spExportSpatial) | Exports a sf object to a spatial layer.
[spMakeSpatialPoints()](#spMakeSpatialPoints) | Generates an S4 SpatialPoints object from X/Y coordinates.
[spReprojectSpatial()](#spReprojectSpatial) | Reprojects a sf object.
[spClipPoint()](#spClipPoint) | Subset points with a Spatial polygon layer.
[spClipPoly()](#spClipPoly) | Subset Spatial polygons layer with another Spatial polygons layer.
[spClipRast()](#spClipRast) | Subset raster layer with a Spatial polygons layer.
[spExtractPoly()](#spExtractPoly) | Extracts point attribute values from Spatial polygons layer(s).
[spExtractRast()](#spExtractRast) | Extracts point attribute values from raster layer(s).
[spGetXY()](#spGetXY) | Wrapper: Extracts XY coordinates and subsets to boundary.
[spGetPlots()](#spGetXY) | Wrapper: Extracts plot data and subsets to boundary.
[spGetAuxiliary()](#spGetAuxiliary) | Wrapper: Extracts point attribute values, area for estimation unit(s), and zonal statistics for strata predictor layers.
[spGetEstUnit()](#spGetEstUnit) | Wrapper: Extracts point attribute values and area for estimation unit(s).
[spGetStrata()](#spGetStrata) | Wrapper: Extracts point attribute values, area for estimation unit(s), and pixel counts for strata spatial layer.
[spUnionPoly()](#spUnionPoly) | Generates one Spatial polygons object from two Spatial polygons layers.
[spZonalRast()](#spZonalRast) | Extracts summary statistics by polygon (i.e., zone) for a raster.


## Objective of tutorial
The objective of this tutorial is to demonstrate how to use `FIESTA`'s spatial tools for manipulating and summarizing spatial data. The examples use data from three inventory years of field measurements in the state of Wyoming, from FIADB_1.7.2.00, last updated June 20, 2018, downloaded on June 25, 2018 and stored as internal data objects in `FIESTA`.

## Example data - Wyoming (WY), Inventory Years 2011-2012

Data Frame | Description
-----------| --------------------------------------------------------------------------------
WYplt      | WY plot-level data
WYcond     | WY condition-level data
WYtree     | WY tree-level data

External data            | Description
-------------------------| ------------------------------------------------------------------
WYbighorn_adminbnd.shp   | Polygon shapefile of WY Bighorn National Forest Administrative boundary*
WYbighorn_districtbnd.shp| Polygon shapefile of WY Bighorn National Forest District boundaries**
WYbighorn_forest_nonforest_250m.tif| GeoTIFF raster of predicted forest/nonforest (1/0) for stratification*** 
WYbighorn_dem_250m.img   | Erdas Imagine raster of elevation change, in meters****


*USDA Forest Service, Automated Lands Program (ALP). 2018. S_USA.AdministrativeForest (\url{http://data.fs.usda.gov/geodata/edw}). Description: An area encompassing all the National Forest System lands administered by an administrative unit. The area encompasses private lands, other governmental agency lands, and may contain National Forest System lands within the proclaimed boundaries of another administrative unit. All National Forest System lands fall within one and only one Administrative Forest Area.

**USDA Forest Service, Automated Lands Program (ALP). 2018. S_USA.RangerDistrict (http://data.fs.usda.gov/geodata/edw). Description: A depiction of the boundary that encompasses a Ranger District.

***Based on MODIS-based classified map resampled from 250m to 500m resolution and reclassified from 3 to 2 classes: 1:forest; 2:nonforest. Projected in Albers Conical Equal Area, Datum NAD27 (Ruefenacht et al. 2008). Clipped to extent of WYbighorn_adminbnd.shp.

****USGS National Elevation Dataset (NED), resampled from 30m resolution to 250m. Projected in Albers Conical Equal Area, Datum NAD27 (U.S. Geological Survey 2017). Clipped to boundary of WYbighorn_adminbnd.shp.



### Set up

First, you'll need to load the `FIESTA` library:

```{r, warning = F, message = F}
library(FIESTA)
```

Next, you'll have to load some external data from the `FIESTA` package and set up objects in your `R` global environment:

```{r}
# File names for external spatial data
WYbhfn <- system.file("extdata",
                      "sp_data/WYbighorn_adminbnd.shp",
                      package = "FIESTA")

WYbhdistfn <- system.file("extdata",
                          "sp_data/WYbighorn_districtbnd.shp",
                          package = "FIESTA")

WYbhdist.att <- "DISTRICTNA"

fornffn <- system.file("extdata",
                       "sp_data/WYbighorn_forest_nonforest_250m.tif",
                       package = "FIESTA")

demfn <- system.file("extdata",
                     "sp_data/WYbighorn_dem_250m.img",
                     package = "FIESTA")

# County-level boundaries for USA and subset for Wyoming
WYco <- stunitco[stunitco$STATENM == "Wyoming",]
```

Finally, you'll need to set up an "outfolder". This is just a file path to a folder where you'd like `FIESTA` to send your data output. For this vignette, we have set our outfolder file path as a temporary directory. .

```{r, include = F}
outfolder <- tempdir()
```

## Examples

In the remainder of this vignette, we provide examples of using the functions in the `sp` portion of `FIESTA`. 

### <a name="spImportSpatial"/>`spImportSpatial()`

The `spImportSpatial` function imports a spatial layer (e.g., ESRI Shapefile) to a simple feature (`sf`) object.

<details>
  <summary>View Example</summary>

```{r}
# Import external data shapefiles
WYbh <- spImportSpatial(WYbhfn)
WYbhdist <- spImportSpatial(WYbhdistfn)

# Display boundary
plot(sf::st_geometry(WYbhdist), border="blue")
plot(sf::st_geometry(WYbh), add=TRUE)

```

</details>

### <a name="spExportSpatial"/>`spExportSpatial()`

The `spExportSpatial` function exports an `sf` object.

<details>
  <summary>View Example</summary>
  
```{r}
# Export Spatial Polygons layer to a shapefile
spExportSpatial(sfobj = WYbh, 
                savedata_opts = savedata_options(out_dsn = "WYbh.shp", 
                                                 outfolder = outfolder, 
                                                 overwrite_dsn = TRUE))
```

</details>

### <a name="spMakeSpatialPoints"/>`spMakeSpatialPoints()`

The `spMakeSpatialPoints` function generates an `sf` points object with a defined projection. Note: The coordinate reference system (crs) is: `prj = "longlat"` and `datum = "NAD83"`.

<details>
  <summary>View Example</summary>

We can use EPSG code with `spMakeSpatialPoints`.

```{r}        
WYspplt <- spMakeSpatialPoints(xyplt = WYplt, 
                               xy.uniqueid = "CN", 
                               xvar = "LON_PUBLIC", 
                               yvar = "LAT_PUBLIC", 
                               xy.crs = 4269)

```        

We can look at the results to verify that WYspplt is now an `sf` points object with the desired projection:
        
```{r}        
WYspplt[ ,c("CN", "geometry")]
```

And finally we export the spatial points to an outfolder. 

```{r}
WYspplt <- spMakeSpatialPoints(xyplt = WYplt,
                               xy.uniqueid = "CN", 
                               xvar = "LON_PUBLIC", 
                               yvar = "LAT_PUBLIC",
                               xy.crs = 4269,
                               exportsp = TRUE,
                               savedata_opts = savedata_options(out_dsn = "spplt",
                                                                out_fmt = "shp",
                                                                outfolder = outfolder,
                                                                out_layer = "WYplots",
                                                                overwrite_layer = TRUE))
```

</details>

### <a name="spReprojectVector"/>`spReprojectVector()`

The `spReprojectVector` function re-projects a spatial vector layer. Note: The layer must have a defined coordinate reference system. 

<details>
  <summary>View Example</summary>
  
```{r}
prj <- "+proj=utm +zone=12 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
WYspplt.utm12 <- spReprojectVector(layer = WYspplt, 
                                   crs.new = prj)
        
# Check results
sf::st_crs(WYspplt.utm12)
```

</details>

### <a name="spClipPoint"/>`spClipPoint()`

The `spClipPoint` function subsets `SpatialPoints` or point file with a Spatial polygon boundary.

<details>
  <summary>View Example</summary>

```{r}
# Get points within Bighorn National Forest boundary (project on the fly)

WYbhptslst <- spClipPoint(xyplt = WYplt, 
                          uniqueid = "CN", 
                          clippolyv = WYbh, 
                          spMakeSpatial_opts = list(xvar = "LON_PUBLIC", 
                                                    yvar = "LAT_PUBLIC",
                                                    xy.crs = 4269))

WYbhptslst <- spClipPoint(xyplt = WYplt, 
                          uniqueid = "CN", 
                          clippolyv = WYbh, 
                          savedata = TRUE, 
                          exportsp = TRUE, 
                          spMakeSpatial_opts = list(xvar = "LON_PUBLIC", 
                                                    yvar = "LAT_PUBLIC",
                                                    xy.crs = 4269), 
                          savedata_opts = list(outfolder = outfolder, 
                                               out_layer = "WYbh"))

names(WYbhptslst)
WYbhspplt <- WYbhptslst$clip_xyplt
WYbhprj <- WYbhptslst$clip_polyv

```

Next we check and display the output:

```{r}
plot(sf::st_geometry(WYbhprj), border="red", lwd=2)
plot(sf::st_geometry(WYbhspplt), add=TRUE)
```

Note: If the projection of spplt is not the same as the points layer, the points layer will be reprojected to the same projection as `clippolyv` (See notes in help file for more info).

Now we generate a sf object first, then clip `sf` points with the `spClipPoint` function:

```{r}
WYspplt <- spMakeSpatialPoints(xyplt = WYplt, 
                               xy.uniqueid = "CN", 
                               xvar = "LON_PUBLIC", 
                               yvar = "LAT_PUBLIC", 
                               xy.crs = 4269)

WYbhptslst <- spClipPoint(xyplt = WYspplt, 
                          uniqueid = "CN",
                          clippolyv = WYbh)
```

We can also subset other tables with clipped points. 

```{r}
WYbhptslst <- spClipPoint(xyplt = WYspplt, 
                          uniqueid = "CN", 
                          clippolyv = WYbh, 
                          othertabnms = c("WYcond", "WYtree"))

names(WYbhptslst)

```

Next, we can export clipped points.

```{r} 
## Export clipped points
spExportSpatial(WYbhspplt, 
                savedata_opts = savedata_options(out_layer = "WYbhpts", 
                                                 outfolder = outfolder, 
                                                 overwrite_layer = TRUE))
```

Finally, we can get points within Bighorn National Forest boundary (project on the fly) and save to an outfolder. 

```{r}
WYbhptslst <- spClipPoint(xyplt = WYplt, 
                          uniqueid = "CN", 
                          clippolyv = WYbh, 
                          othertabnms = c("WYcond", "WYtree"), 
                          exportsp = TRUE, 
                          spMakeSpatial_opts = list(xvar = "LON_PUBLIC", 
                                                    yvar = "LAT_PUBLIC", 
                                                    xy.crs = 4269), 
                          savedata_opts = list(outfolder = outfolder, 
                                               overwrite_layer = TRUE,
                                               outfn.pre = "clip"))

names(WYbhptslst)
```

</details>

### <a name="spClipPoly"/>`spClipPoly()`

The `spClipPoly` function clips (intersects) a polygon vector layer with another polygon vector layer.
<details>
  <summary>View Example</summary>
  
We first subset the WYco `SpatialPolygons` layer with `WYbighorn`. Note: you must download USAco from the `geodata` package and subset to Wyoming. See the "Set Up" section above.
  
```{r}
WYbhco <- spClipPoly(polyv = WYco,
                     clippolyv = WYbh)
```

We can now check and display `WYbighorn_co` with labels.

```{r}
plot(sf::st_geometry(WYbhco['COUNTYCD']), col = sf::sf.colors(nrow(WYbhco)))
coords <- sf::st_coordinates(sf::st_centroid(sf::st_geometry(WYbhco)))
text(coords[,"X"], coords[,"Y"], WYbhco[["COUNTYCD"]])
```

</details>


### <a name="spClipRast"/>`spClipRast()`

The `spClipRast` function subsets a raster to a polygon extent or boundary.

<details>
  <summary>View Example</summary>

First, we subset the strata layer with the Medicine Wheel district boundary

```{r}
WYbhMW <- WYbhdist[WYbhdist$DISTRICTNA == "Medicine Wheel Ranger District",]

plot(sf::st_geometry(WYbhdist))
plot(sf::st_geometry(WYbhMW), border="red", add=TRUE)
```

Next, we can clip the raster. Note: If the projection of `polyv` is not the same as `rast`, `polyv` will by reprojected to the same projection as `rast` before clipping (See note in the help file for more details).

```{r}
WYbhMW.fornf <- spClipRast(rast = fornffn, 
                           clippolyv = WYbhMW, 
                           outfolder = outfolder)
```

Finally, we display the results:

```{r}
WYbhMWprj <- FIESTAutils::crsCompare(WYbhMW, 
                        FIESTAutils::rasterInfo(WYbhMW.fornf)$crs)$x

terra::plot(terra::rast(WYbhMW.fornf))
plot(sf::st_geometry(WYbhMWprj), border = "red", add = TRUE)
```

</details>

### <a name="spExtractPoly"/>`spExtractPoly()`

The `spExtractPoly` function subsets a `SpatialPolygons` layer with another `SpatialPolygons` layer.

<details>
  <summary>View Example</summary>
  
First, we extract polygon attributes from `WYbighorn` to `WYpts`, keeping `NULL` values. Note: If the projection of `spplt` is not the same as the `SpatialPoints`, the `SpatialPoints` layer will be reprojected to the same projection as `clippolyv` before the evaluation points (See the note in help file for more details).

```{r}
extpolylst <- spExtractPoly(WYspplt, 
                            xy.uniqueid = "CN", 
                            polyvlst = WYbhdist)

WYspplt_bh <- extpolylst$spxyext

head(WYspplt_bh)
plot(WYspplt_bh["DISTRICTNA"], pch = 8)
```

Next we can extract a subset of polygon attributes from `WYbighorn` to `WYpts`, not keeping `NULL` values.

```{r}
extpolylst <- spExtractPoly(WYspplt, 
                            xy.uniqueid = "CN", 
                            polyvlst = WYbh, 
                            polyvarlst = c("FORESTNUMB", "FORESTNAME"), 
                            keepNA = FALSE)

WYspplt_bh2 <- extpolylst$spxyext

head(WYspplt_bh2)
```

</details>

### <a name="spExtractRast"/>`spExtractRast()`

The `spExtractRast` function extracts values from one or more raster layers and appends to a SpatialPoints layer or data frame.

<details>
  <summary>View Example</summary>
  
We can extract raster values from `WYdem` to `WYpts`, not keeping `NULL` values.
  
```{r}
extrastlst <- spExtractRast(WYspplt, 
                            rastlst = c(fornffn, demfn), 
                            xy.uniqueid = "CN", 
                            keepNA = FALSE)

WYspplt_dem <- extrastlst$sppltext

```

We can verify that the values from the raster layers were indeed added to our SpatialPoints layer.

``` {r}
head(WYspplt_dem[ , c("WYbighorn_forest_nonforest_250m", "WYbighorn_dem_250m")])
```


</details>

### <a name="spGetAuxiliary"/>`spGetAuxiliary()`

The `spGetAuxiliary` function extracts data and zonal statistics for model-assisted or model-based (small area) estimation. The major steps are as follows:

  1) Check parameters    
  2) Extract point values from `dunit_layer`    
  3) Set up output data structures    
  4) Extract point values and get zonal statistics from continuous raster layers    
  5) Extract point values and get zonal statistics from categorical raster layers    
  6) Get total acres from `domlayer` (if `areacalc = TRUE`)
  
<details>
  <summary>View Example</summary>

First we generate an `sf` object from the WY plot data, public coordinates (`xvar = "LON_PUBLIC"`, `yvar = "LAT_PUBLIC"`). Note: the public coordinate projection information is: `prj = "longlat"`, `datum = "NAD83"`.
  
```{r}
WYspplt <- spMakeSpatialPoints(xyplt = WYplt, 
                               xy.uniqueid = "CN", 
                               xvar = "LON_PUBLIC", 
                               yvar = "LAT_PUBLIC", 
                               xy.crs = 4269)
```

Next, we derive new layers from `dem`. 

```{r}
library(terra)

dem <- rast(demfn)
slpfn <- paste0(outfolder, "/WYbh_slp.img")
slp <- terra::terrain(dem,
                      v = "slope",
                      unit = "degrees",
                      filename = slpfn, 
                      overwrite = TRUE,
                      NAflag = -99999.0)

aspfn <- paste0(outfolder, "/WYbh_asp.img")
asp <- terra::terrain(dem,
                      v = "aspect",
                      unit = "degrees", 
                      filename = aspfn,
                      overwrite = TRUE,
                      NAflag = -99999.0)
```

Finally, we call `spGetAuxiliary` with out points object, boundary layer, and raster files.

```{r}
rastlst.cont <- c(demfn, slp, asp)
rastlst.cont.name <- c("dem", "slp", "asp")
rastlst.cat <- fornffn
rastlst.cat.name <- "fornf"

modeldat <- spGetAuxiliary(xyplt = WYspplt, 
                           uniqueid = "CN", 
                           unit_layer = WYbhfn, 
                           unitvar = NULL, 
                           rastlst.cont = rastlst.cont, 
                           rastlst.cont.name = rastlst.cont.name, 
                           rastlst.cat = rastlst.cat, 
                           rastlst.cat.name = rastlst.cat.name, 
                           rastlst.cont.stat = "mean", 
                           asptransform = TRUE, 
                           rast.asp = asp, 
                           keepNA = FALSE, 
                           showext = FALSE, 
                           savedata = FALSE)
```

We start by getting a sense for all that `spGetAuxiliary` outputs by looking at the names of the `modeldat` object:

``` {r}
names(modeldat)
```

The `pltassgn` object returned by this function contains all of the necessary plot identifiers as well as the extracted values from each of the auxiliary rasters.

``` {r}
head(modeldat$pltassgn)
```

`unitzonal` contains zonal statistics for each domain in the input layer. In this case we are treating the entire layer as a single domain so we just get one row of output.

```{r}
head(modeldat$unitzonal)
```

`unitarea` contains area values by domain in the input layer.

```{r}
head(modeldat$unitarea)
```

Finally, we can look at the resulting predictor names (prednames). 

``` {r}
prednames <- modeldat$prednames
```

</details>

### <a name="spGetXY"/>`spGetXY()`

The `spGetXY` function extracts XY data within a given boundary.

<details>
  <summary>View Example</summary>
  
We can extract public coordinates within WY Bighorn NF boundary, returning spatial (spxy) and nonspatial plot identifiers (pltids).

```{r}
WYbhxy <- spGetXY(bnd = WYbhfn,
                  xy_datsource = "datamart",
                  eval = "FIA",
                  eval_opts = eval_options(Cur = TRUE),
                  returnxy = TRUE)
names(WYbhxy)

pltids <- WYbhxy$pltids
head(pltids)

spxy <- WYbhxy$spxy
plot(sf::st_geometry(WYbhprj))
plot(sf::st_geometry(spxy), add=TRUE)
```


Or, returning only nonspatial plot identifiers (pltids).

```{r}
WYbhxyids <- spGetXY(bnd = WYbhfn,
                     xy_datsource = "datamart",
                     eval = "FIA",
                     eval_opts = eval_options(Cur = TRUE),
                     returnxy = FALSE)
names(WYbhxyids)

pltids <- WYbhxyids$pltids
head(pltids)
```


</details>

### <a name="spGetPlots"/>`spGetPlots()`

The `spGetPlots` function extracts plot data within a given boundary.

<details>
  <summary>View Example</summary>
  
We can extract plots using the public coordinates within WY Bighorn NF boundary. In this example we utilize the FIA datamart to pull all of the necessary tables in and draw plots from the most recent FIA evaluation.

```{r}
WYbhdat <- spGetPlots(bnd = WYbhfn, 
                      states = "Wyoming",
                      datsource = "datamart", 
                      eval = "FIA",
                      eval_opts = eval_options(Cur = TRUE),
                      istree = FALSE)
names(WYbhdat)
```


</details>

### <a name="spGetEstUnit"/>`spGetEstUnit()`

The `spGetEstUnit` function extracts estimation unit values to plots and calculate area by estimation unit.

<details>
  <summary>View Example</summary>
  
First, we create a `SpatialPoints` object from `WYplt`

```{r}
WYspplt <- spMakeSpatialPoints(xyplt = WYplt, 
                               xy.uniqueid = "CN", 
                               xvar = "LON_PUBLIC", 
                               yvar = "LAT_PUBLIC", 
                               xy.crs = 4269)
xyplt <- WYspplt
```



#### By Bighorn National Forest District

We now get acres and strata information for Bighorn National Forest using the districts as our estimation units.

```{r}
unitdat.bhdist <- spGetEstUnit(xyplt = WYplt, 
                               uniqueid = "CN", 
                               unit_layer = WYbhdistfn, 
                               unitvar = "DISTRICTNA", 
                               spMakeSpatial_opts = list(xvar = "LON_PUBLIC", 
                                                         yvar = "LAT_PUBLIC", 
                                                         xy.crs = 4269))

head(unitdat.bhdist$unitarea)
head(unitdat.bhdist$pltassgn[ ,c("CN", "DISTRICTNA")])
```

</details>


### <a name="spGetStrata"/>`spGetStrata()`

The `spGetStrata` function is a wrapper to extract attribute and area from a polygon or raster estimation unit layer and a polygon or raster layer with strata pixel categories. 

<details>
  <summary>View Example</summary>
  
First, we generate an `sf` object from the WY plot data, public coordinates (`xvar = "LON_PUBLIC"`, `yvar = "LAT_PUBLIC"`). Note: The public coordinate projection information is: `prj = "longlat"`, `datum = "NAD83"`.

```{r}
WYspplt <- spMakeSpatialPoints(xyplt = WYplt, 
                               xy.uniqueid = "CN", 
                               xvar = "LON_PUBLIC", 
                               yvar = "LAT_PUBLIC", 
                               xy.crs = 4269)
```

Next, we extract polygon attributes from `WYbighorn` to `WYpts`, keeping `NULL` values.

```{r}
stratlst <- spGetStrata(WYspplt, 
                        uniqueid = "CN", 
                        unit_layer = WYbhfn, 
                        strattype = "RASTER", 
                        strat_layer = fornffn) 
names(stratlst) 
stratlst$stratalut
```


</details>

### <a name="spUnionPoly"/>`spUnionPoly()`

The `spUnionPoly` function generates a unioned `sf` object with polygons and attributes from two `sf` polygon objects.

<details>
  <summary>View Example</summary>
  
First, we import spatial data with `spImportSpatial`, and then extract polygon attributes from `WYbighorn` to `WYpts`, keeping `NULL` values.
  
```{r}
WYbh <- spImportSpatial(WYbhfn)

polyUnion <- spUnionPoly(polyv1 = stunitco[stunitco$STATENM == "Wyoming", ], 
                         polyv2 = WYbh, 
                         areacalc = TRUE)

plot(sf::st_geometry(polyUnion))
head(polyUnion)
```

</details>

### <a name="spZonalRast"/>`spZonalRast()`

The `spZonalRast` function extracts summary statistics by polygon (i.e., zone).

<details>
  <summary>View Example</summary>
  
First, we import spatial data with `spImportSpatial`, and then extract polygon attributes from `WYbighorn` to `WYpts`, keeping `NULL` values.

```{r}
WYbhdist <- spImportSpatial(WYbhdistfn)

zonallst <- spZonalRast(polyv = WYbhdist, 
                        polyv.att = "DISTRICTNA", 
                        rastfn = demfn, 
                        zonalstat = c("mean", "sum")) 
names(zonallst) 

zonalext <- zonallst$zonalext
head(zonalext)
```

</details>

